Untitled

### 2016-07-13 inst/examples/20160715-slides-Krakow.Rmd
Slides almost ready

---
title: "Printing in a Periphery: a Quantitative Study of Finnish Knowledge Production, 1640-1828"
author: "Mikko Tolonen, Jani Marjanen, Hege Roivainen, Leo Lahti"
date: "`r Sys.Date()`"
output: 
  beamer_presentation:
    theme: "boxes"
    colortheme: "orchid"
    fonttheme: "professionalfonts"
---

@@

library(ggmap)
library(gisfin)
library(stringr)
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
library(bibliographica)
library(fennica)
library(sorvi)
library(devtools)
# install_github("ropengov/gisfin")
# devtools::install_github('cttobin/ggthemr')
library(gisfin)
#library(ggthemr)

@@

### TODO

sivu 27. Dissertation languages: kategoria "other" sisällön selvittäminen!

Yleisesti eri kuvien tekstit yms. tsekkaaminen, korostukset yms.

@@

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=3, echo=FALSE}
library(dplyr)
library(ggplot2)
theme_set(theme_bw(20))
df <- df0 %>% filter(publication_year >=1640 & publication_year <=1828) %>% group_by(catalog, publication_decade) %>% tally()
p <- ggplot(df, aes(x = publication_decade, y = n, color = catalog)) + 
       geom_point() + 
       geom_line() + 
       ylab("Documents (n)") + xlab("Publication year") +
       scale_color_manual(values = c("blue", "darkgreen")) + 
       guides(fill = guide_legend(reverse = TRUE, title = "")) #+ 
print(p)
```

@@

**Collaboration & Reuse** further use and enrichment

@@

**1. General trends & political events**

@@

### Turku as a Swedish University Town among Uppsala, Tartu and Stockholm

```{r map1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=20, fig.height=5, eval=TRUE}
library(dplyr)
library(ggmap)
library(ggplot2)
library(gisfin)
source("funcs.R")
theme_set(get_theme_map())
mapdata <- filter(df0,
          (catalog == "Kungliga" &
	      publication_place %in% c("Stockholm", "Uppsala",
	                         "Greifswald", "Lund", "Tartu"))|
	  (catalog == "Fennica" &
	      publication_place == "Turku")) %>%
      group_by(publication_decade, publication_place,
	       latitude, longitude) %>%
      summarise(n = n())

@@

region <- "Europe.north"
mymap <- get_map(location=geobox(region), color = "bw",
                  source="google",  maptype="terrain")

p0 <- ggmap(mymap) + theme(legend.position="none")

pics <- list()
for (i in c(1650, 1710, 1760, 1800)) {

    # Plot map
    p <- p0

    # Pick the investigated period (sliding window)
    # Sum up the years within this sliding window for each element    	    
    dfw <- subset(mapdata, publication_decade == i) %>%
    	     group_by(publication_place, latitude, longitude, highlight) %>%
		       summarize(n = sum(size))

    if (nrow(dfw) > 0) {
      p <- p0 +
      	geom_point(data = dfw,
      	  aes(x = longitude, y = latitude,
	      size = 20*log10(1+n),
	      color = highlight)) +
	      #color = highlight), alpha = 0.8) +	      
	  ggtitle(i) +
	  scale_size(range = c(2, 15)) +
	  theme(title = element_text(size = 30)) +
	  scale_color_manual(values = c("black", "red"))	  
    }

  pics[[as.character(i)]] <- p

}

library(gridExtra)
grid.arrange(pics[[1]], nrow = 1)
```

@@

- 1700-1721 Fennica drop
- 1760-1770s Kungliga peak 
- 1803-1815 Napoleonic wars

```{r publishingovertime, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4, out.width="200px"}
library(magrittr)
library(ggplot2)
library(dplyr)
df2 <- df0 %>% group_by(publication_decade, catalog) %>%
               summarise(n = n())
df2$highlight <- rep(FALSE, nrow(df2))
df2$highlight[df2$catalog == "Kungliga" &
	      df2$publication_decade %in% c(1770, 1810)] <- TRUE
df2$highlight[df2$catalog == "Fennica" &
	      df2$publication_decade %in% c(1710, 1810)] <- TRUE
v <- seq(1500, 1800, 50)
theme_set(theme_bw(20))
p <- ggplot(df2, aes(x = publication_decade, y = n)) +
     geom_line(aes(color = catalog), size = 1) +
     geom_point(aes(shape = catalog, color = highlight, size = highlight)) +
     ylab("Title count (n)") + xlab("Publication year") +
     #theme(legend.position = "none") +
     scale_x_continuous(breaks = v, labels = v) +
     scale_color_manual(values = c("black", "blue", "darkgreen", "blue")) + 
     guides(fill = guide_legend(reverse = TRUE)) 
print(p)
```

### Swedish 1760-1770 publishing peak

Pamphleteering plays a crucial role during abolished censorship

```{r pamflets1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4, out.width="280px", fig.show="hold"}
df <- df0 %>%
        filter(catalog == "Kungliga")

# Add indicator of the publication period
df$period <- cut(df$publication_year,
           breaks = c(1757, 1766, 1775, 1783),
		 include.lowest = TRUE,
		 right = FALSE
		 )

@@

p <- ggplot(df,
       aes(x = period, y = paper.per.title)) + 
       geom_bar(stat = "identity") +
       scale_x_discrete(labels = c("1757-1765", "1766-1774", "1775-1783")) +
       ylab("Paper per title") + xlab("") 
print(p)
```

@@

  rect_left <- c(min(na.omit(dfm$date)),
               1719-.5, 1719+.5, # Stockholm 20 januari 1719 1 juni 1719
               1734-.5, 1734+.5, # Stockholm 14 maj 1734 14 december 1734
               1765-.5, 1766+.5, # Stockholm 21 februari 1765 21 oktober 1766
               1769-.5, 1770+.5, # Norrköping & Stockholm 22 april 1769 5 februari 1770
               1771-.5, 1772+.5, # Stockholm 19 juni 1771 12 september 1772
               1792-.5, 1792+.5, # Gävle 26 januari 1792 24 februari 1792         
               max(na.omit(dfm$date)))
  rectangles <- data.frame(
    xmin = rect_left[-length(rect_left)],
    xmax = rect_left[-1],
    ymin = ymin,
    ymax = ymax
    )
  rectangles$shade <- rep(c("White", "Highlight"), length = nrow(rectangles))

 riksplace = c("Stockholm", "Stockholm", "Stockholm", "Norrköping", "Stockholm", "Gävle")
 cols = c("gray", "lightblue", "yellow", "white")
 rectangles$shade[rectangles$shade == "Highlight"] = riksplace
 rectangles$shade = factor(rectangles$shade)

@@

**2. Publishers in Sweden and Turku**

@@

@@

### Top publishers in Turku/Fennica (title count)

```{r publishers4-fennica, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=7, fig.width=15}
place <- "Turku"
catalogue = "Fennica"
df = df0

@@

# Title count per decade & publisher	
npub <- df %>% group_by(publication_decade, publisher) %>% tally()

# TITLE COUNT
theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = n)) +
       geom_bar(stat = "identity", position = "stack", aes(fill = publisher)) + 
       xlab("Publication year") +
       ylab("Title count (n)") #+        
       #ggtitle(paste("Title count per publisher (", place, "/", catalogue, ")", sep = ""))
print(p)
```

@@

**3. Languages & religious texts in Finland**

@@

p <- ggplot(df, aes(x = publication_year, y = n, color = language)) +
       geom_line() + geom_point()

@@

### Devotional literature (catechisms, hymns, prayer collections, etc.)

Form of literature denoting the advancement of reading in Finland in the nineteenth century. It has been unclear when does it show up as a relevant category.

@@

- play an important role in early modern knowledge production. It has not been analysed in what language they were in Finland (and Sweden).

@@

**4. University curriculum & Enlightenment in Finland**

@@

```{r disslang, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=6}
catalogue <- "Fennica"
langs <- c("Finnish", "Swedish", "Latin", "German",
           "Russian", "French", "Hebrew", "Other")

@@

```{r topics-103B, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=9}
sel = c("filosofia","luonnonfilosofia","metafysiikka","logiikka")
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
                 paper = sum(paper.consumption.km2, na.rm = TRUE))

# PAPER CONSUMPTION
p1 <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_bar(aes(fill = language), position = "stack", stat = "identity") +                     
       xlab("Publication year") +
       ylab("Paper consumption") +
       guides(fill = guide_legend(reverse = TRUE, title = "")) +
       ggtitle("Philosophy")

@@

# PAPER CONSUMPTION
p2 <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_bar(aes(fill = language), position = "stack", stat = "identity") +             
       xlab("Publication year") +
       ylab("Paper consumption") +
       guides(fill = guide_legend(reverse = TRUE, title = "")) +
       ggtitle("History")

@@

# PAPER CONSUMPTION
theme_set(theme_bw(20))
p3 <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_bar(aes(fill = language), position = "stack", stat = "identity") +              
       xlab("Publication year") +
       ylab("Paper consumption") +
       guides(fill = guide_legend(reverse = TRUE, title = "")) +
       ggtitle("Economy")
grid.arrange(p1, p2, p3, nrow = 3)
```

@@

Paper consumption for different document formats over time.

```{r LIBER-13, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=11, fig.height=9, fig.show="hold", out.width="160px"}
for (catal in unique(df0$catalog)) {
  df2 <- subset(df0, catalog == catal) %>% group_by(publication_decade, gatherings) %>% summarize(paper.consumption.km2 = sum(paper.consumption.km2, na.rm = TRUE), n = n()) 
  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 15)), "NA"))
  df2$highlight <- rep("gray", nrow(df2))
  df2$highlight[df2$gatherings == "8vo"] <- "darkgreen"

  p <- ggplot(df2, aes(y = paper.consumption.km2,
                       x = publication_decade,
  	       shape = gatherings,
		       linetype = gatherings))		       
  p <- p + geom_point(size = 4)
  p <- p + geom_smooth(method = "loess", size = 1,
         aes(color = highlight, fill = highlight))
  p <- p + scale_color_manual(values = c("darkgreen", "darkgray"))
  p <- p + scale_fill_manual(values = c("darkgreen", "darkgray"))  	 
  p <- p + ggtitle("Paper consumption in time by gatherings")
  p <- p + xlab("Year")
  p <- p + ylab("Paper consumption (km2)")
  p <- p + guides(linetype = guide_legend(keywidth = 5),
       	          shape = guide_legend(keywidth = 5))
  p <- p + ggtitle(paste("Paper consumption\n(", catal, ")", sep = ""))
  print(p)
}
```

@@

For more info, see: \textcolor{blue}{https://github.com/rOpenGov/fennica}

>>>>

---
title: "Printing in a Periphery: a Quantitative Study of Finnish Knowledge Production, 1640-1828"
author: "Mikko Tolonen, Jani Marjanen, Hege Roivainen, Leo Lahti"
date: "`r Sys.Date()`"
output: 
  beamer_presentation:
    theme: "boxes"
    colortheme: "orchid"
    fonttheme: "professionalfonts"
fontsize: 13pt
---

@@

library(ggmap)
library(gisfin)
library(stringr)
library(dplyr)
library(tidyr)
library(reshape2)
library(ggplot2)
library(bibliographica)
library(fennica)
library(sorvi)
library(devtools)
# install_github("ropengov/gisfin")
# devtools::install_github('cttobin/ggthemr')
#library(ggthemr)

@@

@@

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=3, echo=FALSE}
library(dplyr)
library(ggplot2)
theme_set(theme_bw(20))
df <- df0 %>% filter(publication_year >=1640 & publication_year <=1828) %>% group_by(catalog, publication_decade) %>% tally()
p <- ggplot(df, aes(x = publication_decade, y = n, color = catalog)) + 
       geom_point() + 
       geom_line() + 
       ylab("Documents (n)") + xlab("Publication year") +
       scale_color_manual(values = c("blue", "darkgreen")) + 
       guides(color = guide_legend(reverse = TRUE, title = "")) #+ 
print(p)
```

@@

@@

\Large{1. General trends and political events}

@@

### Turku as a Swedish University Town among Uppsala, Tartu and Stockholm (+ Lund & Greifswald)

```{r map1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=20, fig.height=4, eval=TRUE}
library(dplyr)
library(ggmap)
library(ggplot2)
library(gisfin)
source("funcs.R")
theme_set(get_theme_map())
i <- 1650
mapdata1 <- filter(df0,
              publication_decade == i &
              (catalog == "Kungliga" & 
	            publication_place %in% c("Stockholm", "Uppsala",
	                         "Greifswald", "Lund", "Tartu"))|
	  (catalog == "Fennica" &
	      publication_place == "Turku")) %>%
      group_by(publication_decade, publication_place,
	       latitude, longitude) %>%
      summarise(n = n())

mapdata <- bind_rows(mapdata1, 
  as.data.frame(list(publication_decade = 1650, publication_place = "Lund", 
                     latitude = 55.70293, longitude = 13.19294, n = 0)))

@@

mapdata$highlight2 <- rep(1, nrow(mapdata))
mapdata$highlight2[mapdata$publication_place %in% c("Greifswald", "Lund")] <- 9
mapdata$highlight2 <- factor(mapdata$highlight2)

region <- c(8, 40, 31, 76)
names(region) <- c("left", "bottom", "right", "top")
#mymap <- get_map(location=geobox(region), color = "bw", source="osm",  maptype="terrain")
#mymap <- get_map(location=geobox(region), source="stamen", maptype = "watercolor")
mymap <- get_map(location=geobox(region), maptype="roadmap")

# Pick the investigated period (sliding window)
# Sum up the years within this sliding window for each element    	    
dfw <- mapdata %>%
    	     group_by(publication_place, latitude, longitude, highlight, highlight2) %>%
		       summarize(n = sum(size))
dfw$size <- log10(1 + dfw$n)
dfw$size[!dfw$highlight2 == 1] <- max(dfw$size)

# Plot map
p <- ggmap(mymap) + 
      	geom_point(data = dfw,
      	  aes(
      	  x = longitude, 
      	  y = latitude,
	        size = size,
	        shape = highlight2, 
	        color = highlight)) +
	        ggtitle(i) +
	        scale_size(range = c(2, 15)) +
	        theme(title = element_text(size = 30), legend.position="none") +
	        scale_color_manual(values = c("black", "red")) +
	        scale_shape_manual(values = c(19,1))
print(p)
```

Map by Stamen

@@

- 1700-1721 Fennica drop
- 1760-1770 Pamphleteering plays a crucial role during abolished censorship
- 1803-1815 Napoleonic wars

```{r publishingovertime, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=4, out.width="300px"}
library(magrittr)
library(ggplot2)
library(dplyr)
df2 <- df0 %>% group_by(publication_decade, catalog) %>%
               summarise(n = n())
df2$highlight <- rep(FALSE, nrow(df2))
df2$highlight[df2$catalog == "Kungliga" &
	      df2$publication_decade %in% c(1770, 1810)] <- TRUE
df2$highlight[df2$catalog == "Fennica" &
	      df2$publication_decade %in% c(1710, 1810)] <- TRUE
v <- seq(1500, 1800, 100)
theme_set(theme_bw(20))
p1 <- ggplot(df2, aes(x = publication_decade, y = n)) +
     geom_line(aes(color = catalog), size = 1) +
     geom_point(aes(shape = catalog, color = catalog, size = highlight)) +
     ylab("Title count (n)") + xlab("Publication year") +
     #theme(legend.position = "none") +
     scale_x_continuous(breaks = v, labels = v) +
     scale_color_manual(values = c("blue", "darkgreen")) + 
  	 scale_size_discrete(range = c(2, 8)) +
     guides(color = guide_legend(reverse = TRUE), size = "none", shape = "none")

df <- df0 %>% filter(catalog == "Kungliga")
# Add indicator of the publication period
df$period <- cut(df$publication_year,
           breaks = c(1757, 1766, 1775, 1783),
		 include.lowest = TRUE,
		 right = FALSE
		 )

@@

p2 <- ggplot(df,
       aes(x = period, y = paper.per.title)) + 
       geom_bar(stat = "identity") +
       scale_x_discrete(labels = c("1757-1765", "1766-1774", "1775-1783")) +
       ylab("Paper per title") + xlab("")

library(gridExtra)
grid.arrange(p1, p2, nrow = 1)
```

@@

  rect_left <- c(min(na.omit(dfm$date)),
               1719-.5, 1719+.5, # Stockholm 20 januari 1719 1 juni 1719
               1734-.5, 1734+.5, # Stockholm 14 maj 1734 14 december 1734
               1765-.5, 1766+.5, # Stockholm 21 februari 1765 21 oktober 1766
               1769-.5, 1770+.5, # Norrköping & Stockholm 22 april 1769 5 februari 1770
               1771-.5, 1772+.5, # Stockholm 19 juni 1771 12 september 1772
               1792-.5, 1792+.5, # Gävle 26 januari 1792 24 februari 1792         
               max(na.omit(dfm$date)))
  rectangles <- data.frame(
    xmin = rect_left[-length(rect_left)],
    xmax = rect_left[-1],
    ymin = ymin,
    ymax = ymax
    )
  rectangles$shade <- rep(c("Background", "Highlight"), length = nrow(rectangles))

 riksplace = c("Stockholm", "Stockholm", "Stockholm", "Norrköping", "Stockholm", "Gävle")
 cols = c("white", "black", "blue", "darkgray")
 rectangles$shade[rectangles$shade == "Highlight"] = riksplace
 rectangles$shade = factor(rectangles$shade)

@@

\Large{2. Publishers in Sweden and Turku}

@@

\begin{center}
[Publishers in Sweden and Turku]
\end{center}

@@

### Top publishers in Turku/Fennica

```{r publishers4-fennica, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, fig.height=6, fig.width=12}
library(dplyr)
place <- "Turku"
catalogue = "Fennica"
df = df0

@@

# Title count per decade & publisher	
npub <- df %>% group_by(publication_decade, publisher) %>% tally() %>% arrange(n)
npub$publisher <- factor(npub$publisher, levels = (npub %>% group_by(publisher) %>% summarise(total = sum(n)) %>% arrange(total))$publisher)

# TITLE COUNT
theme_set(theme_bw(20))
p <- ggplot(npub, aes(x = publication_decade, y = publisher)) +
      geom_point(aes(size = n)) + 
      #geom_line(aes(color = publisher), size = 2) +
      #geom_smooth(aes(fill = publisher)) +
      # geom_bar(stat = "identity", position = "stack", aes(fill = publisher)) + 
       xlab("Year") +
       ylab("") +
       guides(size = guide_legend(title = "Title count (n)"))
       #ggtitle(paste("Title count per publisher (", place, "/", catalogue, ")", sep = ""))
print(p)
```

@@

\Large{3. Languages and religious texts in Finland}

@@

p <- ggplot(df, aes(x = publication_year, y = n, color = language)) +
       geom_line() + geom_point() + ylab("Documents (n)") + xlab("Publication year")

@@

### Devotional literature (catechisms, hymns, prayers, etc.)

Form of literature denoting the advancement of reading in Finland in the 19th century. It has been unclear when does it show up as a relevant category.

@@

- Important role in early modern knowledge production. It has not been analysed in what language they were in Finland (and Sweden).

@@

\Large{4. University curriculum and Enlightenment in Finland}

@@

```{r disslang, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=5}
catalogue <- "Fennica"
langs <- c("Finnish", "Swedish", "Latin", "German",
           "Russian", "French", "Hebrew", "Other")

@@

```{r topics-103B, echo=FALSE, message=FALSE, warning=FALSE, fig.width=12, fig.height=8}
sel = c("filosofia","luonnonfilosofia","metafysiikka","logiikka")
df = dfl
df$hit <- apply((sapply(sel, function (x) {grepl(x, tolower(df$subject_topic))})), 1, any)  
df <- df %>% filter(hit)
df <- df %>% group_by(publication_decade, language) %>%
             summarise(n = n(),
                 paper = sum(paper.consumption.km2, na.rm = TRUE))

# PAPER CONSUMPTION
p1 <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_bar(aes(fill = language), position = "stack", stat = "identity") +                     
       xlab("Publication year") +
       ylab("Paper") +
       guides(fill = guide_legend(reverse = TRUE, title = "")) +
       ggtitle("Philosophy")

@@

# PAPER CONSUMPTION
p2 <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_bar(aes(fill = language), position = "stack", stat = "identity") +             
       xlab("Publication year") +
       ylab("Paper") +
       guides(fill = guide_legend(reverse = TRUE, title = "")) +
       ggtitle("History")

@@

# PAPER CONSUMPTION
theme_set(theme_bw(20))
p3 <- ggplot(df, aes(x = publication_decade, y = paper, group = language)) +
       geom_bar(aes(fill = language), position = "stack", stat = "identity") +              
       xlab("Publication year") +
       ylab("Paper") +
       guides(fill = guide_legend(reverse = TRUE, title = "")) +
       ggtitle("Economy")
grid.arrange(p1, p2, p3, nrow = 3)
```

@@

Paper consumption: different document formats over time

```{r LIBER-13, echo=FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.width=11, fig.height=9, fig.show="hold", out.width="160px"}
for (catal in unique(df0$catalog)) {
  df2 <- subset(df0, catalog == catal) %>% group_by(publication_decade, gatherings) %>% summarize(paper.consumption.km2 = sum(paper.consumption.km2, na.rm = TRUE), n = n()) 
  df2 <- filter(df2, gatherings %in% setdiff(names(which(table(df2$gatherings) >= 15)), "NA"))
  df2$highlight <- rep("Other", nrow(df2))
  df2$highlight[df2$gatherings == "8vo"] <- "Octavo"

  p <- ggplot(df2, aes(y = paper.consumption.km2,
                       x = publication_decade,
  	       shape = gatherings,
		       linetype = gatherings))		       
  p <- p + geom_point(size = 4)
  p <- p + geom_smooth(method = "loess", size = 1,
         aes(color = highlight, fill = highlight))
  p <- p + scale_color_manual(values = c("darkgreen", "darkgray"))
  p <- p + scale_fill_manual(values = c("darkgreen", "darkgray"))  	 
  p <- p + ggtitle("Paper consumption in time by gatherings")
  p <- p + xlab("Year")
  p <- p + ylab("Paper consumption (km2)")
  p <- p + guides(linetype = guide_legend(keywidth = 5),
       	          shape = guide_legend(keywidth = 5))
  p <- p + ggtitle(catal)
  print(p)
}
```

@@

For more info, see:

- \textcolor{blue}{https://github.com/rOpenGov/fennica}

### 2016-07-13 inst/examples/slides.Rmd
Slides almost ready

>>>>

[General trends & political events]

### Publishing activity over time

- 1700-1721 Fennica drop
- 1760-1770 Kungliga peak 
- 1803-1815 Napoleonic wars

```{r publishingovertime, echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=4, out.width="200px"}
library(magrittr)
library(ggplot2)
library(dplyr)
df2 <- df0 %>% group_by(publication_decade, catalog) %>%
               summarise(n = n())
df2$highlight <- rep(FALSE, nrow(df2))
df2$highlight[df2$catalog == "Kungliga" &
	      df2$publication_decade %in% c(1770, 1810)] <- TRUE
df2$highlight[df2$catalog == "Fennica" &
	      df2$publication_decade %in% c(1710, 1810)] <- TRUE
v <- seq(1500, 1800, 100)
theme_set(theme_bw(20))
p <- ggplot(df2, aes(x = publication_decade, y = n)) +
     geom_line(aes(color = catalog), size = 1) +
     geom_point(aes(shape = catalog, color = catalog, size = highlight)) +
     ylab("Title count (n)") + xlab("Publication year") +
     #theme(legend.position = "none") +
     scale_x_continuous(breaks = v, labels = v) +
     scale_color_manual(values = c("blue", "darkgreen")) + 
  	 scale_size_discrete(range = c(2, 8)) +
     guides(color = guide_legend(reverse = TRUE), size = "none", shape = "none") 
print(p)
```

[General trends & political events]

### Swedish 1760-1770 publishing peak

Pamphleteering plays a crucial role during abolished censorship

```{r pamflets1, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4, out.width="280px", fig.show="hold"}
df <- df0 %>% filter(catalog == "Kungliga")

# Add indicator of the publication period
df$period <- cut(df$publication_year,
           breaks = c(1757, 1766, 1775, 1783),
		 include.lowest = TRUE,
		 right = FALSE
		 )

# Remove entries outside the investigated publication periods
df <- df %>% filter(!is.na(period)) %>%
	     group_by(period) %>%
	     summarise(n = n(),
		  paper = sum(paper.consumption.km2, na.rm = TRUE)) %>%
        mutate(paper.per.title = paper/n)

p <- ggplot(df,
       aes(x = period, y = paper.per.title)) + 
       geom_bar(stat = "identity") +
       scale_x_discrete(labels = c("1757-1765", "1766-1774", "1775-1783")) +
       ylab("Paper per title") + xlab("") 
print(p)
```
